|

### 2023-04-16 paper.Rmd
conclusion tweaking and paragraph on changing sig levels

Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. (Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.) The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05. It is also interesting to note that setting the significance level much lower, .001, .0001, .00001, ... for the RESET test, so that it rejects less does shift the power curve towards that of the visual test. However, it never reaches the matches the performance of the visual test.

For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests. The visual test power curve is a right shift. This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test. From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5. However the BP test would reject at around effect size 2.5. Interestingly, the power curve for the SW test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too. The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.

@@

@@

The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans. That is, a conventional test conclude there are problems with the model fit almost twice as often as a human would. They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.

@@

Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots. A residual plot may contains many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions. These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, results in a set of more accurate visual findings. This enables a careful calibration for reading structure in residual plots.

However, human evaluation of residuals is expensive. It is time-consuming, laborious and unfriendly to vision-impaired people. This is another reason why it often appears to be ignored. With the availability of sophisticated computer vision algorithms today, the goal of this work is to form the basis of providing automated residual plot reading. The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable. We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with. Details about this system will be discussed in our next paper.

The experiment also revealed some interesting details about how residual plots are read. For the most part, the visual test performed very similarly to the appropriate conventional test only with the power curve shifted in the less sensitive direction. Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals.

As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test. Structure is hardest to detect if fitted values are lognormal. Also, complex structure are generally harder to detect, but there are outliers. Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions. A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.

>>>>

Figure \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. (Note that the BP and SW tests have much lower power, which is expected because they are not designed to detect non-linearity.) The bootstrapped power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject at a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25, for a significance level of 0.05.

For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests. The visual test power curve is a right shift. This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test. From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5. However the BP test would reject at around effect size 2.5. Interestingly, the power curve for the SW test (for non-normality) is only slightly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too. The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.

@@

Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) it would be curious to know whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test. Interestingly, it doesn't. Despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (See Appendix).

@@

The experiment found that conventional residual-based statistical tests are more sensitive to weak departures from model assumptions than visual tests as would be evaluated by humans. That is, a conventional test concludes there are problems with the model fit almost twice as often as a human would. They often reject when departures in the form of non-linearity and heteroskedasticity are not visibly different from null residual plots.

@@

Now it is important to note that residual plots need to be delivered as a lineup, where it is embedded in a field of null plots. A residual plot may contain many visual features, but some are caused by the characteristics of the predictors and the randomness of the error, not by the violation of the model assumptions. These irrelevant visual features have a chance to be filtered out by subjects with a comparison to null plots, resulting in more accurate reading. The lineup enables a careful calibration for reading structure in residual plots.

Human evaluation of residuals is expensive, though. It is time-consuming and laborious. This is possibly why residual plot analysis is often not done, in practice. However, with the emergence of effective computer vision this work could form the basis to provide automated residual plot reading. <!--The findings suggest the strong demand of graphical inspection in regression diagnostics, so developing an automatic visual inference system to evaluate lineups of residual plots is valuable. We plan to build a completed open-source system in an `R` package and provide a web interface such as a website for public to interact with. Details about this system will be discussed in our next paper.-->

The experiment also revealed some interesting results. For the most part, the visual test performed very similarly to the appropriate conventional test except that it has a power curve shifted in the less sensitive direction. Unlike the conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals. If the fitted value distribution is not uniform, there is a small loss of power in the visual test. But quite a surprise to us was the big difference in power of the conventional test, which appear to be quite sensitive to how the predictors are sampled. Another surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.
